---
title: "Berry Report"
author: "Yuxin Zeng"
date: "2020/10/19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,warning=F,message=F)
pacman::p_load("knitr","tidyverse","magrittr","stringr","ggplot2")
```

###1.Introduction
The NASS website has a large amount of berries data from many states in recent years, but these data are very messy, with mixed variables and different dimensions. After some simple cleaning and organization, I chose blueberry for EDA and present it in a shiny app.

###2.Data Peparation
First of all, read in the data and remove those columns with single repeated values which are meaningless. "State ANSI" is also removed from the dataset since the "State" is clear enough. We find that berries data had only 8 out of 21 columns containing meaningful data. But some of the variables in these 8 columns are not clear enough, and we need to deal with them later.
When it comes to "Period" column."Year" generally refers to calendar year while the definition of "Marketing year" varies by commodity. For Prices Received data, they both refers to an unweighted average (by month). Only consider the "YEAR" period. 
Filter out berries with specific numbers in "Value" in order to do further analysis. 
Divide the data into three parts according to the type of berries, so we get blueberries dataset, raspberries dataset, and strawberries dataset. 
```{r}
#Read the data
berries=read_csv("berries.csv",col_names=T)

#Remove columns with single value
col=berries%>%summarize_all(n_distinct)
single=which(col[1,]==1)
berries%<>%select(-all_of(single))
#Remove State ANSI
berries%<>%select(-4)
#Period="Year"
berries=berries%>%filter(Period=="YEAR")
#Filter out berries with specific numbers in "Value"
berries%<>%filter(Value!="(D)")
berries%<>%filter(Value!="(NA)")

#Group the data by commodity
bberry=berries%>%filter(Commodity=="BLUEBERRIES")
rberry=berries%>%filter(Commodity=="RASPBERRIES")
sberry=berries%>%filter(Commodity=="STRAWBERRIES")
```

Focus on blueberries. Separate multiple variables from the same column, merge variables of the same type into a new column, and delete duplicate columns.
```{r}
#Separate "Data Item"
bberry%<>%separate('Data Item',c("B","type","meas","what"), sep=",") 
bberry%<>%separate(type,c("b1","type","b2","lab1", "lab2"),sep="")
bberry%<>%mutate(label=paste(lab1,lab2)) 
bberry%<>%select(-c(B,b1,b2)) 

##Domain & Domain Category
bberry%<>%separate(Domain,c("D_left","D_right"),sep=",")
bberry%<>%mutate(D_left="CHEMICAL",D_left="") 
bberry%<>%mutate(Chemical=paste(D_left,D_right)) 
bberry%<>%select(-c(D_left,D_right)) 

bberry%<>%separate('Domain Category',c("DC_left","DC_right"), sep=",")
bberry%<>%separate(DC_left,c("DC_left_l","DC_left_r"),sep=":")
bberry%<>%separate(DC_right,c("DC_right_l","DC_right_r"),sep=":")
bberry%<>%select(-c(DC_left_l,DC_right_l))

bberry%<>%select(Year,State,what,meas,label,Chemical,DC_left_r,DC_right_r,Value)
```

Some variables are not properly separated. We have entries in both the "what" and "meas" columns that begin with "MEASURED IN". Separate them from their current columns and then merge them to unit column.
```{r}
#Write a function
f1 <- function(a,b){
  if(a){
    return(b)
  }else{
    return("")
  }
}

f1_log=c(F,T,T)
f1_str=c("one","two","three")
map2(f1_log,f1_str,f1)

#Replace "NA" with blank before using the function
bberry[is.na(bberry)]=""

#"Meas"
detect.meas=str_detect(bberry$meas,"MEASURED IN")
bberry%<>%mutate(new_col1=unlist(map2(detect.meas,bberry$meas,f1))) 
bberry%<>%mutate(meas=str_replace(bberry$meas,"MEASURED IN.*$", ""))

#"What"
detect.what=str_detect(bberry$what,"MEASURED IN")
bberry%<>%mutate(new_col2=unlist(map2(detect.what,bberry$what,f1))) 
bberry%<>%mutate(what=str_replace(bberry$what,"MEASURED IN.*$", ""))

#Units
bberry%<>%mutate(units=str_trim(paste(new_col1,new_col2)))
```

Finally organize the columns and rename them.
```{r}
#Rename the columns
bberry%<>%rename(Marketing=meas,Avg=what,Harvest=label,Chem_family=DC_left_r,Materials=DC_right_r,Measures=units)
#Joint some columns
bberry%<>%mutate(production=str_trim(paste(Marketing,Harvest)))
bberry%<>%mutate(Chemical=str_trim(paste(Chem_family,Chemical)))

bberry%<>%select(Year,State,production,Avg,Measures,Materials,Chemical,Value)
write.csv(bberry,file="C:/Users/lenovo/Desktop/615 R/berry/bberry.csv")
```

###3.EDA
Values are measured in different way, for example, some are measured in dollars, some are measured in LB. Only choose the blueberries that are measured in LB. Explore and visualize the relationship between Values and Year, State, production.
```{r}
#Summary
options(scipen=200)
bberry$Value=as.numeric(gsub(",","",bberry$Value))
dim(bberry) 
summary(bberry)
p=qplot(x=Value,data=bberry,ylab='frequency')
p

#Choose ine measurement
sum(bberry$Measures=="MEASURED IN LB")
sum(bberry$Measures=="MEASURED IN $")
sum(bberry$Measures=="MEASURED IN LB / ACRE / APPLICATION")
sum(bberry$Measures=="MEASURED IN LB / ACRE / YEAR")
bberry%<>%filter(Measures=="MEASURED IN LB")

#Year
p1=boxplot(Value~Year,data=bberry,xlab='year',ylab='value')
p1

#State
p2=ggplot(bberry,aes(x=Year,y=Value,fill=State))+geom_bar(position="dodge",stat="identity")+xlab("year")+ylab("value")+labs(fill="State")
p2

#Production
p3=ggplot(bberry,aes(x=Year,y=Value,fill=production))+geom_bar(position="dodge",stat="identity")+xlab("year")+ylab("value")+labs(fill="production")
p3
```

###4.Reference
<div id="refer-anchor-1"></div>
[1] [National Agricultural Statistics Service](https://quickstats.nass.usda.gov/)
<div id="refer-anchor-2"></div>
[2] [Visit Maine](https://visitmaine.com/things-to-do/agricultural-attractions/wild-for-blueberries)
<div id="refer-anchor-3"></div>
[3] [Vince Vu](https://github.com/vqv/ggbiplot)
